Sequential Composition for Relaxed Memory: Pomsets with Predicate Transformers

### 2021-07-02 paper/drf.tex
messing with c11 to make drf easier

\section{Local Data Race Freedom and Sequential Consistency}
\label{sec:sc}

Idea: merging reads can only make a difference is if there is a race.
Let $\sempomca{i}{}$ be the semantic function generated by applying the
construction of \reflem{lem:po} to $\semmca{i}{}$.

We adapt \citeauthor{Dolan:2018:BDR:3192366.3192421}'s
[\citeyear{Dolan:2018:BDR:3192366.3192421}] notion of \emph{Local Data Race
  Freedom (LDRF)} to our setting.

The result requires that locations are properly initialized.  We assume a
sufficient condition: that programs have the form
``$\aLoc_1\GETS\aVal_1\SEMI \cdots \aLoc_n\GETS\aVal_n\SEMI\aCmd$'' where
every location mentioned in $\aCmd$ is some $\aLoc_i$.  To simplify the
definition of \emph{happens-before}, we ban fences and \RMWs.
% To simplify
% the proof, we assume there are no local declarations of the form
% $(\VAR\aLoc\SEMI \aCmd)$.

@@

\bookmark{Definitions}
\paragraph{Data Race}
Data races are defined using \emph{program} order $(\rpox)$, not
\emph{pomset} order $(\le)$. %, and thus is stable with respect to augmentation.
In \ref{SB}, for example, $(\DR{x}{0})$ has an $x$-race with $(\DW{x}{1})$,
but not $(\DW{x}{0})$, which is $\rpox$-before it.

It is obvious how to enhance the semantics of prefixing and most other
operators to define $\rpox$.  When combining pomsets using the conditional,
the obvious definition may result in cycles, since $\rpox$-ordered reads may
coalesce---see the discussion of \ref{CA} in \textsection\ref{sec:refine}.  In
this case we include a separate pomset for each way of breaking these cycles.

Because we ignore the features of
\textsection\ref{sec:variants}, we can adopt the simplest definition of
\emph{synchronizes\hyp{}with}~($\rsw$): Let $\bEv\xsw\aEv$ exactly when
$\bEv$ fulfills $\aEv$, $\bEv$ is a release, $\aEv$ is an acquire, and
$\lnot(\bEv\xpox\aEv)$.

Let ${\rhb}=({\rpox}\cup{\rsw})^+$ be the \emph{happens-before} relation.  In
\ref{Pub1}, for example, $(\DW{x}{1})$ happens-before $(\DR{x}{0})$, but this
fails if either $\mRA$ access is relaxed.

Let $L\subseteq\Loc$ be a set of locations.  We say that $\bEv$ \emph{has an
  $L$-race with} $\aEv$ (notation $\bEv\lrace{L}\aEv$) when at least one is
relaxed, they \emph{conflict}
(Def.~\ref{def:rf}) at
some location in $L$, and they are unordered by $\rhb$: neither $\bEv\xhb\aEv$ nor
$\aEv\xhb\bEv$.

@@

>>>>

% \section{Local Data Race Freedom and Sequential Consistency}
\section{LDRF-SC for \PwTmcaTITLE{}}
\label{sec:sc}

\cbstart
In this appendix, we establish a \drfsc{} for \PwTmca{2}.  We prove an
\emph{external} result, where the notion of \emph{data-race} is independent
of the semantics itself.  Since every \PwTmca{2} is also a \PwTmca{1}, the
result also applies there.  Our result is also \emph{local}.  Using
\citeauthor{Dolan:2018:BDR:3192366.3192421}'s
[\citeyear{Dolan:2018:BDR:3192366.3192421}] notion of \emph{Local Data Race
  Freedom (LDRF)}.

We do not address \PwTc{}.  The existing \drfsc{} result for \cXI{}
\cite{DBLP:phd/ethos/Batty15} does not rely on dependencies and thus applies
to \PwTc{}.  However, \citeauthor{DBLP:phd/ethos/Batty15}'s result is
\emph{internal}, where data-races are defined using the semantics of \cXI{}
itself.  Using the notion of dependency defined here, it should be possible
to prove an stronger external result for \cXI{}---we leave this as future
work.

\citet{DBLP:journals/pacmpl/JagadeesanJR20} prove \ldrfsc{} for Pomsets with
Preconditions (\PwP{}).  \PwTmca{} generalizes \PwP{} to account for
sequential composition.  Most of the machinery of \ldrfsc{}, however, has
little to do with sequential semantics.  Thus, we have borrowed heavily from
the text of \cite{DBLP:journals/pacmpl/JagadeesanJR20}; indeed, we have
copied directly from the \LaTeX{} source, which is publicly available.  We
indicate substantial changes or additions using a change-bar on the right.

There are two main changes:
\begin{itemize}
\item \PwP{} assumes \ref{pom-rf-le} ($\rrfx$ implies $\le$)
  and thus is similar to \PwTmca{1}.  \PwTmca{2} is a weaker model that is
  new to this paper.  Hence we cannot use
  \citeauthor{DBLP:journals/pacmpl/JagadeesanJR20}'s result directly.
\item \citeauthor{DBLP:journals/pacmpl/JagadeesanJR20} did not provide an
  accurate account of program order for merged actions.  We correct this
  deficiency.  Let $\sempomca{i}{}$ be the semantic function generated by
  applying the construction of \reflem{lem:po} to $\semmca{i}{}$.  The proof
  concerns only \emph{complete} pomsets.
\end{itemize}

\cbend

The result requires that locations are properly initialized.  We assume a
sufficient condition: that programs have the form
``$\aLoc_1\GETS\aVal_1\SEMI \cdots \aLoc_n\GETS\aVal_n\SEMI\aCmd$'' where
every location mentioned in $\aCmd$ is some $\aLoc_i$.  To simplify the
definition of \emph{happens-before}, we ban fences and \RMWs.

@@

\bookmark{Definitions}
\paragraph{Data Race}
Data races are defined using \emph{program} order $(\rpox)$, not
\emph{pomset} order $(\le)$. %, and thus is stable with respect to augmentation.

Because we ban fences and \RMWs, we can adopt the simplest definition of
\emph{synchronizes\hyp{}with}~($\rsw$): Let $\bEv\xsw\aEv$ exactly when
$\bEv$ fulfills $\aEv$, $\bEv$ is a release, $\aEv$ is an acquire, and
$\lnot(\bEv\xpox\aEv)$.

Let ${\rhb}=({\rpox}\cup{\rsw})^+$ be the \emph{happens-before} relation.

Let $L\subseteq\Loc$ be a set of locations.  We say that $\bEv$ \emph{has an
  $L$-race with} $\aEv$ (notation $\bEv\lrace{L}\aEv$) when
\begin{enumerate*}
\item at least one is relaxed, 
\item at least one is a write,
\item they access the same location in $L$, and
\item they are unordered by $\rhb$: neither $\bEv\xhb\aEv$ nor
  $\aEv\xhb\bEv$.
\end{enumerate*}

@@

  In the above argument, it is unimportant whether $\aEv$ reads-from an
  internal or an external write; thus the argument applies to
  \PwTmca{2} and \PwTmca{1} as it does for \PwTmca{1}.
